A Classical Density-Functional Theory for Describing Water Interfaces

### 2012-08-20 papers/water-SAFT/paper.tex

A number of classical density functionals have been developed for
water~\cite{ding1987, Yang1992, yang1994density, gloor2002saft, 
  gloor2004accurate, gloor2007prediction, Jaqaman2004, chuev2006, clark2006developing,
  lischner2010classical, fu2005vapor-liquid-dft,kiselev2006new,
  blas2001examination}, each of which
captures some of the qualitative behavior of water.  However, each of
these functionals also fail to capture some of water's unique
properties.  For instance, the functional of Lischner \emph{et
  al}\cite{lischner2010classical} treats the surface tension
correctly, but can only be used at room temperature, and thus captures
none of the temperature-dependence of water.  A number of classical
density functionals have recently been produced that are based on
Statistical Associating Fluid Theory (SAFT)\cite{ 
  yu2002fmt-dft-inhomogeneous-associating,
  fu2005vapor-liquid-dft,gloor2002saft,muller2001molecular,
  clark2006developing, gloor2007prediction, gloor2004accurate,
  gross2009density, kahl2008modified, blas2001examination}.  These
functionals are based on a perturbative thermodynamic expansion, and
do reproduce the temperature-dependence of water's properties.

@@

Walther \emph{et al}\cite{walther2004hydrophobic} studied the
interactions between two carbon nanotubes, which are geometrically
similar to our hydrophobic rods, using molecular dynamics with the SPC
model for water, and a Lennard-Jones potential for the interaction of
carbon with water, for nanotubes of diameter 1.25~nm and separations
ranging from about 0.3~nm to 1.5~nm.  The SPC model underestimates the
surface tension of water by about 24\%\cite{vega2007surface}, so we
cannot expect this work to provide quantitative agreement with real
water.  Walther \emph{et al} observe a drying transition between the
two nanotubes, which occurs at a smaller radius than our results
suggest.  However, when Walther \emph{et al} disable the attractive
interaction between nanotube and water, the drying effect occurs at
much longer range, in agreement with our results.  \fixme{Please read
  this carefully, as I did some major changes.  I took a look at the
  paper, and it seems that in Figure 3c (and the paragraph that ends
  at the top of that page) they finally do a truly hydrophobic
  interaction, which is a better comparison with what we did.  The
  drying transition is inhibited by the attraction between tube and
  water. }

>>>>

A number of classical density functionals have been developed for
water~\cite{ding1987, Yang1992, yang1994density, gloor2002saft,
  gloor2004accurate, gloor2007prediction, Jaqaman2004, chuev2006,
  clark2006developing, chuev2006,
  lischner2010classical, fu2005vapor-liquid-dft,kiselev2006new,
  blas2001examination}, each of which
captures some of the qualitative behavior of water.  However, each of
these functionals also fail to capture some of water's unique
properties.  For instance, the functional of Lischner \emph{et
  al}\cite{lischner2010classical} treats the surface tension
correctly, but can only be used at room temperature, and thus captures
none of the temperature-dependence of water.  A functional by Chuev
and Skolov\cite{chuev2006} uses an ad hoc modification of FMT that can
predict hydrophobic hydration near temperatures of 298~K, but does not produce a
correct equation of state due to their method producting a high value
for pressure. A number of classical density functionals have recently been produced that are based on
Statistical Associating Fluid Theory (SAFT)\cite{ 
  yu2002fmt-dft-inhomogeneous-associating,
  fu2005vapor-liquid-dft,gloor2002saft,muller2001molecular,
  clark2006developing, gloor2007prediction, gloor2004accurate,
  gross2009density, kahl2008modified, blas2001examination}.  These
functionals are based on a perturbative thermodynamic expansion, and
do reproduce the temperature-dependence of water's properties.

@@

Walther \emph{et al}\cite{walther2004hydrophobic} studied the
interactions between two carbon nanotubes, which are geometrically
similar to our hydrophobic rods, using molecular dynamics with the SPC
model for water, and a Lennard-Jones potential for the interaction of
carbon with water, for nanotubes of diameter 1.25~nm and separations
ranging from about 0.3~nm to 1.5~nm.  The SPC model underestimates the
surface tension of water by about 24\%\cite{vega2007surface}, so we
cannot expect this work to provide quantitative agreement with real
water.  Walther \emph{et al} observe a drying transition between the
two nanotubes, which occurs at a smaller radius than our results
suggest.  However, when Walther \emph{et al} disable the attractive
interaction between nanotube and water, the drying effect occurs at
much longer range, in agreement with our results.
